Challenging Fuel Cycle Modeling Assumptions: Facility and Time Step Discretization Effects

### 2015-10-20 paper.tex
clean up @gonuke's edits

    \item Groups of facilities (usually all facilities of a given type) are
        treated as a single, aggregate entity. Therefore, all facilities experience
        identical operational histories.

    \item Resource flows are modeled as continuous, within the limitations of time
        step discretization.  Therefore, material is acquired and discharge in
        uniform amounts at every time step.

@@

    \item Reactor capacity is deployed and retired in increments of single
      reactor units, each with capable of producing power $P_r$ (e.g. 1 GWe
      LWRs).  $N_{r}$ reactors are deployed at any time capable of producing
      $P_r \cdot N_r$ power.

@@

\begin{table}[h]
    \caption{Reactor Facility Parameters}
    \centering
    \begin{tabular}{ |r | c c | }
        \hline                       
            & LWR & SFR \\
        \hline                       
        Lifetime (yr)         & 80 & 80 \\
        Cycle length (months) & 18 & 14 \\  %% why list this SFR cycle length if it's never used?
        Batch size (kg)       & 29565 & 7490 \\
        Batches per core      & 3 & 5 \\
        \hline                       
    \end{tabular}
    \captionsetup{justification=centering}
    \label{tab:reactors}
\end{table}

The described scenario is used as closely as possible in each of the four
cases.  However, a few small case-dependent changes to the scenarios relating to
reactor configuration are necessary.  For case \emph{QI}, 
cycle and outage times must be adjusted to be a
multiple of 3 months. 
%% this tweak is a distraction.  Why not simply declare the SFR cycle length &
%% outage as 12/3 from the beginning.  Is this because the fuel recipes you
%% are using assume the other parameters and you don't want to sweep that
%% under the rug?
 In order to make the comparison more valid, the cases
with monthly time steps are also adjusted to have identical cycle and outage
times.  The invariants preserved with respect to reactor behavior between all
four cases are shown in Table \ref{tab:invar}.  Table \ref{tab:reactor-detail} shows
the computed/selected configuration for both LWR and SFR reactor types for all
four cases.

@@

\begin{table*}
    \centering
    \captionsetup{justification=centering}
    \caption{Reactor Parameters by Case}
    \begin{tabular}{ |r | c c | c c | }
        \hline                       
                                          & \multicolumn{2}{c |}{LWR}       & \multicolumn{2}{c |}{SFR} \\
                                          & Cases MI,QI & Cases MF,QF (fleet) & Cases MI,QI & Cases MF,QF (fleet)  \\
        \hline                       
        Cycle length (months)             & 15        & 18                & 12        & 15 \\
        Refueling outage (months)         & 3         & 0                 & 3         & 0 \\
        Batch size ($\text{kg} \cdot \text{HM}$) & 29,565    & 88,695            & 8,025     & 40,125 \\
        Power capacity (MWe)              & 1,080     & 900               & 450       & 360 \\
        \hline                       
    \end{tabular}
    \label{tab:reactor-detail}
\end{table*}

In each of the four cases, the initial 100 LWRs are modified to have a zero
length refueling outage with an explicit 900 MWe net power output capacity.  Because the
individual reactor model doesn't have any way to start a reactor mid-cycle,
this was done to avoid having all initial LWRs refuel at the same time.  The
parameters for each case were also selected to keep capacity factors for
individual reactors very roughly around 80\%.

@@

One difference between the fleet and individual facility modeling is the
\emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
cores is short one half batch of fuel, it will produce $\frac{2.5}{3}\cdot P_r$
of power for that time step where $P_r$ is the power capacity of a single
reactor of the fleet (\emph{e.g.}~1000~MWe).  If an individually modeled reactor is
short a half batch of fuel, then it will produce no power ($0 \cdot P_r$) that
time step.  Time step duration can amplify this effect.  For both fleet and
individual reactor modeling, power goes offline in full time step increments.
Consider an example of a 3-batch reactor with a fuel shortage of $\frac{1}{2}$ batch of
fuel that would be resolve in 1 month.  Depending on the modeling choice, each reactor
would produce the following amounts of energy over a 3 month period:

\begin{itemize}
  \item case \emph{QF}: $\frac{2.5}{3} \cdot 3 \text{mo} \cdot P_r = 2.5P_r$
  \item case \emph{MF}: $[\frac{2.5}{3} \cdot 1 \text{mo} + 1 \cdot 2
    \text{mo}] \cdot P_r = 2.83P_r$
  \item case \emph{QI}: $0 \cdot 3 \text{mo} \cdot P_r = 0$
  \item case \emph{MI}: $[0\cdot 1 \text{mo} + 1 \cdot 2\text{mo}] \cdot P_r = 2 P_r$
\end{itemize}

@@

Quantifying imperfect fuel sharing is a bit difficult, but Figure
\ref{fig:badshare} provides one way to see the \emph{fuel sharing effect}
showing wasted batch-months computed using Equation \ref{eq:badshare}.  Fleet
reactor modeling, by design, exhibits perfect fuel sharing by design (zero
inefficiency) and so is not included in this figure.  Every batch that is
given to a reactor that ends up not being able to start its cycle at the
scheduled time (due to fuel shortage) adds one to this cumulative for each
time step the cycle start is delayed.

>>>>

    \item Groups of facilities (usually all facilities of a given type) are
        treated as a single, aggregate entity. Therefore, all facilities experience
        a single, identical history.

    \item Resource flows are modeled as continuous, within the limitations of time
        step discretization.  Therefore, material is acquired and discharged in
        uniform amounts at every time step.

@@

    \item Reactor capacity is deployed and retired in increments of single
      reactor units, each capable of producing power $P_r$ (e.g. 1 GWe
      LWRs).  $N_{r}$ reactors are deployed at any time capable of producing
      up to $P_r \cdot N_r$ power.

@@

\begin{table}[h]
    \caption{Reactor Facility Parameters}
    \centering
    \begin{tabular}{ |r | c c | }
        \hline                       
            & LWR & SFR \\
        \hline                       
        Lifetime (yr)         & 80 & 80 \\
        Cycle length (months) & 18 & 15 \\
        Batch size (kg)       & 29565 & 7490 \\
        Batches per core      & 3 & 5 \\
        \hline                       
    \end{tabular}
    \captionsetup{justification=centering}
    \label{tab:reactors}
\end{table}

The described scenario is used as closely as possible in each of the four
cases.  However, a few small case-dependent changes to the scenarios relating to
reactor configuration are necessary.
The invariants preserved with respect to reactor behavior between all
four cases are shown in Table \ref{tab:invar}.  Table \ref{tab:reactor-detail} shows
the computed/selected configuration for both LWR and SFR reactor types for all
four cases.

@@

\begin{table*}
    \centering
    \captionsetup{justification=centering}
    \caption{Selected Reactor Parameters by Case}
    \begin{tabular}{ |r | c c | c c | }
        \hline                       
                                          & \multicolumn{2}{c |}{LWR}       & \multicolumn{2}{c |}{SFR} \\
                                          & Cases MI,QI & Cases MF,QF (fleet) & Cases MI,QI & Cases MF,QF (fleet)  \\
        \hline                       
        Cycle length (months)             & 15        & 18                & 12        & 15 \\
        Refueling outage (months)         & 3         & 0                 & 3         & 0 \\
        Batch size ($\text{kg} \cdot \text{HM}$) & 29,565    & 29,565            & 8,025     & 8,025 \\
        Power capacity (MWe)              & 1,080     & 900               & 450       & 360 \\
        \hline                       
    \end{tabular}
    \label{tab:reactor-detail}
\end{table*}

In each of the four cases, the initial 100 LWRs are modified to have a zero
length refueling outage with an explicit 900 MWe net power output capacity.  Because the
individual reactor model doesn't have any way to start a reactor mid-cycle,
this was done to avoid having all initial LWRs refuel at the same time.  The
parameters for each case were also selected to keep capacity factors for
individual reactors very roughly around 90\%, although they are actually 
closer to 80\% due to integer variable constraints on the case MI and QI
reactor configuration.

@@

One difference between the fleet and individual facility modeling is the
\emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
cores is short one half batch of fuel, it will produce $\frac{2.5}{3}\cdot P_r$
of power for that time step where $P_r$ is the power capacity of a single
reactor of the fleet (\emph{e.g.}~1000~MWe).  If an individually modeled reactor is
short one half batch of fuel, then it will produce no power ($0 \cdot P_r$) that
time step.  Time step duration can amplify this effect.  For both fleet and
individual reactor modeling, power goes offline in full time step increments.
Consider an example of a 3-batch reactor with a fuel shortage of $\frac{1}{2}$ batch of
fuel that would be resolve in 1 month.  Depending on the modeling choice, each reactor
would produce the following amounts of energy over a 3 month period:

\begin{itemize}
  \item case \emph{QF}: $\frac{2.5}{3} \cdot 3 \text{ mo} \cdot P_r = 2.5P_r$
  \item case \emph{MF}: $[\frac{2.5}{3} \cdot 1 \text{ mo} + 1 \cdot 2
    \text{mo}] \cdot P_r = 2.83P_r$
  \item case \emph{QI}: $0 \cdot 3 \text{ mo} \cdot P_r = 0$
  \item case \emph{MI}: $[0\cdot 1 \text{ mo} + 1 \cdot 2\text{mo}] \cdot P_r = 2 P_r$
\end{itemize}

@@

Quantifying imperfect fuel sharing is a bit difficult, but Figure
\ref{fig:badshare} provides one way to see the \emph{fuel sharing effect}
showing wasted batch-months computed using Equation \ref{eq:badshare}.  Fleet
reactor modeling, by design, exhibits perfect fuel sharing (zero
inefficiency) and so is not included in this figure.  Every batch that is
given to a reactor that ends up not being able to start its cycle at the
scheduled time (due to fuel shortage) adds one to this cumulative for each
time step the cycle start is delayed.
